Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-03-14 paper/paper.tex
back to 10 pages

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

The goal of this paper is to advance formal methods in elections by
demonstrating how STV vote counting can be verified in a modular,
transparent, and practical feasible way over a minimal computational
trust base:  we show how to synthesise efficient and provably
correct verifiers for a large class of STV schemes in a modular and
almost fully automated way.

@@

Executable evidence verifiers are constructed almost
automatically:  nearly all of the ensuing proof obligations are discharged generically, and we fully  automate the subsequent translation into machine-level verified executable using CakeML.

@@

We also identify a general algorithmic mechanism for performing
computation on evidence, which realises the \emph{commonalities}
among the members of the STV family. This generic method comprises seven general actions%~(Figure~\ref{GenSteps})  
~(Definition~\ref{transitions}) corresponding to actions that tally officers take when counting votes.
\begin{definition}[Generic Counting Actions]\label{transitions}
 The following comprise the generic STV counting actions.
 %The generically appearing counting actions in the STV family are the followings. 
%\begin{figure}  
 \begin{itemize}
\item\textbf{count} for counting the uncounted ballots,
\item\textbf{elect} for electing all or some of the electable candidates, 
\item\textbf{transfer-elected} for distributing surplus votes of elected, %candidates, 
\item\textbf{eliminate} to exclude one or some candidates, 
\item\textbf{transfer-removed} to distribute votes of   the eliminated, %candidates, 
\item\textbf{elected-win} for terminating tallying whenever the vacancies are filled or there is no continuing candidate left, 
\item\textbf{hopeful-win} for terminating tallying whenever  the number of already elected and continuing candidates collectively does not exceed the initial vacancies. 
\end{itemize}
%\caption{Generic STV Counting Steps}
%\label{GenSteps}
%\end{figure}
\end{definition}
Each action consists of  preconditions on when and how the action
comes into effect, and postconditions dictating what effect applying
the action brings about by describing the new state to which the
computation proceeds. In fact, these conditions are the formal
counterparts of the legal clauses in the textual specification of
the counting scheme which inform tally officers how to proceed with tallying, and they are common between all flavours of STV.

@@

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ 
 where $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
\noindent
Differences exist between counting algorithms of STV schemes.  For
example in the STV used for electing the upper house representatives
of the Australian State of Victoria (Victoria STV)~\cite{vic},
distributing votes of an eliminated candidate occurs step by step in
order of the magnitude of the fractional values that those votes
carry.  Consequently, a pre-condition for applying the action is to
first assure that upon each instance of applying transfer-removed
action (a) exactly those votes which have the same fractional value
are transferred and (b) the transfer happens according to the
magnitude order. In contrast, for example, the STV used in the lower
house elections of the Australian state of Tasmania (Tasmania STV)
does not have such precondition simply because votes of an
eliminated candidate are transferred in a single application of
transfer-removed. There are also variations in postconditions. For
example transfer-elected of the ACT STV~\cite{act} used in the lower
house elections of the Australian Capital Territory  requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

\begin{definition}[verifier]\label{verifier}
Assume $\hat{\mathcal{A}}= \langle \mathcal{S}, \mathcal{T},
(S'_t)_{t \in \mathcal{T}} \rangle$ is an instantiation of the STV
algorithm $\mathcal{A}$. A \emph{verifier} for instances of
computation of $\mathcal{A}$ is a function $\hat{\mathcal{V}}$
mapping $\mathbb{Q}\times\mathbb{N}\times 2^{C}\times$
\textsf{List}$(\mathcal{S})$ to $\{0,1\}$ such that for any putative evidence $\hat{\omega} = (Q,s,C,\Omega)$ where $\Omega=\langle\Omega_{1},\dots,\Omega_{n}\rangle$, $\hat{\mathcal{V}}(\hat{\omega}) = 1$ if and only if the following holds:\\
$\forall i\in\{1,\dots,n-1\}.$ 
%\hspace*{0.2cm}$\forall \Omega_{i-1}$ $\Omega_{i}\in\mathcal{S}.$
%$\exists \Omega_{1}$ $\Omega_{2}\in$ \textsf{List} $(\mathcal{S}).$ $\Omega = \Omega_{1} + [\Omega_{i-1};\Omega_{i}] + \Omega_{2}$  
$\exists t\in\mathcal{T}.$  $(\Omega_{i-1},\Omega_{i})\in t$ $\wedge~\vdash S'_{t}[\Omega_{i-1},\Omega_{i}]$\\ 
where $\mathbb{Q}$ and $\mathbb{N}$ respectively represent the set
of rational and natural numbers, $2^{C}$ is the set of all subsets
of $C$, and \textsf{List}$(\mathcal{S})$ is the set of all possible
lists of machine states. The state $\Omega_{i-1}$ is a pre-state and $\Omega_{i}$ is a post-state visited in the execution. The last line means that pre- and post-conditions  of applying $t$ to move from $\Omega_{i-1}$ to $\Omega_{i}$  are logically true. 
\end{definition} 
We shall next provide an example to concretely illustrate a piece of evidence and exemplify how STV schemes work. 
 \subsection{An Example of a Concrete Piece of Evidence}
\label{subsec:InstEv}
\noindent
Figure~\ref{EvInst} shows evidence of a small election whose counting algorithm is the ACT STV 
with three candidates $a$, $b$ and $c$, and initially recorded ballots containing
$b_1=([a,c],1/1)$, $b_2=([a,b,c],1/1)$, $b_3=([a,c,b],1/1)$,
$b_4=([b,a],1/1)$, $b_5=([c,b,a],1/1)$.  The first component of each ballot is a preference-ordered list of candidates, and the second is the fractional transfer value (initially set to $1/1$). The evidence consists of a \emph{header} that specifies the quota (computed according to the ACT STV instructions), the number of vacancies, and the list of competing candidates. 
The fourth line is the election result, and the remainder of the evidence are non-final states visited to compute this
outcome. Each non-final state shows uncounted ballots, tallies of candidates, piles of votes attracted, backlog of the elected, backlog of the eliminated, list of elected candidates, and list of continuing candidates.

@@

\section{The Framework Architecture}\label{sec:Arc} 
\noindent
We progress through three macro-level phases to develop the framework. For the first phase we use HOL4 to formally specify, implement and verify the generic STV, its instantiations, verifiers, and every auxiliary assertions needed for these ends. The second step relies on the verified proof translation tool of CakeML to provably correctly translate implementations of the first phase into equivalent CakeML implementations. The last phase uses the ecosystem of CakeML and its verified compiler to generate machine executable evidence verifiers.

@@

\end{tikzpicture}
\caption{Architecture of the Framework, note that direction of arrow represents module dependencies}
\label{PicArc}
\end{figure}
\paragraph*{\textbf{Compilation}} 
Instantiations of CakeML's  compiler for generating machine executable verifiers happen in this module. For any deeply embedded $\mathcal{V}^{*}_{i}$, corresponding to $\hat{\mathcal{V}}_{dec}$ in \textbf{STV}$_{i}$, using proofs  established in \textbf{DeepSpec}, we instantiate the compiler to synthesise executable  verifier $\mathcal{V}_{i}$.  
\section{Generic STV Formalised}\label{sec:GenCertVer}
We proceed through three steps, namely specification, implementation and verification of implementations in HOL4 to formally verify the machine and its components.
\paragraph*{Specification}  
Recall that we require minimising the trusted computing base  for computing with generated verifiers. The specification module consists of properties which we can use to verify that our implementations of  STV algorithms in HOL4 indeed match with the description of the counting methods used.  
 In light of Definition~\ref{STVMachine} these properties   actually form the semantics of the generic STV machine.   
\paragraph*{Implementation} Our final objective for creating the
framework is to produce means for validating concrete items of evidence. To this end, we need decision procedures used for deciding whether or not a given evidence is valid. Therefore we define boolean-valued functions in the implementation sub-module to computationally realise logical declarations in the specification module. In particular, we implement decision procedures whose computational content  provably realises the specification of the machine semantics.  
\paragraph*{Verification} Here we formally prove that the implementations meet the expectations of their respective specifications. Therefore through verification we come to eliminate a trusted layer  required to lay in our framework.

@@

  \HOLConst{ValidInitCandList}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{distinct}\;\HOLFreeVar{l}
The predicate ValidInitCandList which realises item (6) asserts that each candidate has a pile and a tally. We also declare that elements in the first component of the tally $t$ and pile $p$ are distinct. Therefore we come to satisfy item (1).
  \HOLConst{Valid_PileTally}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{c}.\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{c}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{c}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})
We implement for each of the above declarations a counterpart 
%computational 
\underline{dec}ision procedure that is later translated and then extracted as part of the verifier for actual computation. To illustrate how this phase proceeds, we provide some instances. The following two functions together implement Valid\textunderscore{}PileTally.
  \HOLConst{Valid_PileTally_dec1}\;\HOLConst{\ensuremath{[\,]}}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{true}\\
\HOLConst{Valid_PileTally_dec1}\;(\HOLFreeVar{h}\HOLSymConst{::}\HOLFreeVar{t})\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLConst{\HOLConst{mem}}\;(\HOLConst{fst}\;\HOLFreeVar{h})\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally_dec1}\;\HOLFreeVar{t}\;\HOLFreeVar{l}
  \HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{true}\\
\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;(\HOLFreeVar{l\sb{\mathrm{0}}}\HOLSymConst{::}\HOLFreeVar{ls})\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLKeyword{if}\;\HOLConst{\HOLConst{mem}}\;\HOLFreeVar{l\sb{\mathrm{0}}}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})\;\HOLKeyword{then}\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;\HOLFreeVar{ls}\;
\HOLKeyword{else}\;\HOLConst{false}
We prove that the implementations and specification match:
  \HOLConst{Valid_PileTally}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLConst{Valid_PileTally_dec1}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;\HOLFreeVar{l}
Conjunctions of the computational implementations  define a computational twin TransferAuxDec for the specification  of the transfer-elected semantics TransferAuxSpec. 
  \HOLConst{TransferAuxDec}\;(\HOLFreeVar{qu}\HOLSymConst{,}\HOLFreeVar{st}\HOLSymConst{,}\HOLFreeVar{l})\;\HOLFreeVar{ba}\;\HOLFreeVar{t}\;\HOLFreeVar{t\sp{\prime}}\;\HOLFreeVar{p}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{bl}\;\HOLFreeVar{bl\sb{\mathrm{2}}}\;\HOLFreeVar{bl\sb{\mathrm{2}}\sp{\prime}}\;\HOLFreeVar{e}\;\HOLFreeVar{e\sp{\prime}}\;\HOLFreeVar{h}\;\HOLFreeVar{h\sp{\prime}}\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLConst{null}\;\HOLFreeVar{bl\sb{\mathrm{2}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{e\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{e}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{h\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{t\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{t}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{length}\;\HOLFreeVar{e}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLFreeVar{st}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{list_MEM_dec}\;(\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenDoublePlus}\;\HOLFreeVar{e})\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{list_MEM_dec}\;\HOLFreeVar{bl}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{distinct}\;(\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenDoublePlus}\;\HOLFreeVar{e})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally_dec1}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally_dec1}\;\HOLFreeVar{p}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{p}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}
\;\HOLConst{Valid_PileTally_dec1}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenNeg{}}\HOLConst{null}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{distinct}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{null}\;\HOLFreeVar{bl\sb{\mathrm{2}}\sp{\prime}}\\
\;\;\HOLConst{distinct}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{distinct}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{p})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenNeg{}}\HOLConst{null}\;\HOLFreeVar{bl}\;\HOLSymConst{\HOLTokenConj{}}
\;\HOLConst{null}\;\HOLFreeVar{ba}

>>>>

\setlength{\textfloatsep}{5pt}
\setlength{\intextsep}{5pt}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

The goal of this paper is to advance formal methods in elections by
demonstrating how STV counting can be verified in a modular,
transparent, and practical feasible way over a minimal computational
trust base:  we show how to synthesise efficient and provably
correct verifiers for a large class of STV schemes in a modular and
almost fully automated way.

@@

Executable evidence verifiers are constructed almost
automatically:  nearly all of the ensuing proof obligations are discharged generically, and we fully  automate the translation into machine-level verified executable using CakeML.

@@

We also identify a general algorithmic mechanism to perform
computation on evidence, which realises the \emph{commonalities}
among the members of the STV family. This comprises seven general actions%~(Figure~\ref{GenSteps})  
~(Definition~\ref{transitions}) corresponding to actions that tally officers take when counting votes.
\begin{definition}[Generic Counting Actions]\label{transitions}
 The following comprise the generic STV counting actions.
 %The generically appearing counting actions in the STV family are the followings. 
%\begin{figure}  
 \begin{itemize}
\item\textbf{count} for counting the uncounted ballots,
\item\textbf{elect} for electing all or some of the electable candidates, 
\item\textbf{transfer-elected} for distributing surplus votes of elected, %candidates, 
\item\textbf{eliminate} to exclude one or some candidates, 
\item\textbf{transfer-removed} to distribute votes of   the eliminated, %candidates, 
\item\textbf{elected-win} for terminating tallying whenever the vacancies are filled or there is no continuing candidate left, 
\item\textbf{hopeful-win} for terminating tallying whenever  the number of already elected and continuing candidates collectively does not exceed the initial vacancies. 
\end{itemize}
%\caption{Generic STV Counting Steps}
%\label{GenSteps}
%\end{figure}
\end{definition}
Each action consists of  preconditions on when and how the action
comes into effect, and postconditions dictating what effect applying
the action brings about by describing the new state to which the
computation proceeds. In fact, these conditions are the formal
counterparts of the legal clauses in the textual specification of
the counting scheme which inform tally officers how to proceed with tallying, and they are common between all flavours of STV.

@@

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ 
 where $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
%\noindent
Differences exist between counting algorithms of STV schemes.  For
example in the STV used for electing the upper house representatives
of the Australian State of Victoria (Victoria STV)~\cite{vic},
distributing votes of an eliminated candidate occurs step by step in
order of the magnitude of the fractional values that those votes
carry.  Consequently, a pre-condition for applying the action is to
first assure that upon each instance of applying transfer-removed
action (a) exactly those votes which have the same fractional value
are transferred and (b) the transfer happens according to the
magnitude order. In contrast, for example, the STV used in the lower
house elections of the Australian state of Tasmania (Tasmania STV)
does not have such precondition simply because votes of an
eliminated candidate are transferred in a single application of
transfer-removed. There are also variations in postconditions. For
example transfer-elected of the ACT STV~\cite{act} used in the lower
house elections of the Australian Capital Territory  requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

\begin{definition}[verifier]\label{verifier}
Assume $\hat{\mathcal{A}}= \langle \mathcal{S}, \mathcal{T},
(S'_t)_{t \in \mathcal{T}} \rangle$ is an instantiation of the STV
algorithm $\mathcal{A}$. A \emph{verifier} for instances of
computation of $\mathcal{A}$ is a function $\hat{\mathcal{V}}$
mapping $\mathbb{Q}\times\mathbb{N}\times 2^{C}\times$
\textsf{List}$(\mathcal{S})$ to $\{0,1\}$ such that for any putative evidence $\hat{\omega} = (Q,s,C,\Omega)$ where $\Omega=\langle\Omega_{1},\dots,\Omega_{n}\rangle$, $\hat{\mathcal{V}}(\hat{\omega}) = 1$ if and only if the following holds:\\
$\forall i\in\{1,\dots,n-1\}.$ 
%\hspace*{0.2cm}$\forall \Omega_{i-1}$ $\Omega_{i}\in\mathcal{S}.$
%$\exists \Omega_{1}$ $\Omega_{2}\in$ \textsf{List} $(\mathcal{S}).$ $\Omega = \Omega_{1} + [\Omega_{i-1};\Omega_{i}] + \Omega_{2}$  
$\exists t\in\mathcal{T}.$  $(\Omega_{i-1},\Omega_{i})\in t$ $\wedge~\vdash S'_{t}[\Omega_{i-1},\Omega_{i}]$\\ 
where $\mathbb{Q}$ and $\mathbb{N}$ respectively represent the set
of rational and natural numbers, $2^{C}$ is the set of all subsets
of $C$, and \textsf{List}$(\mathcal{S})$ is the set of all possible
lists of machine states. The state $\Omega_{i-1}$ is a pre-state and $\Omega_{i}$ is a post-state visited in the execution. The last line means that pre- and post-conditions  of applying $t$ to move from $\Omega_{i-1}$ to $\Omega_{i}$  are logically true. 
\end{definition} 
We shall next provide an example to concretely illustrate a piece of evidence and exemplify how STV schemes work. 
 \subsection{An Example of a Concrete Piece of Evidence}
\label{subsec:InstEv}
%\noindent
Figure~\ref{EvInst} shows evidence of a small election whose counting algorithm is the ACT STV 
with three candidates $a$, $b$ and $c$, and initially recorded ballots containing
$b_1=([a,c],1/1)$, $b_2=([a,b,c],1/1)$, $b_3=([a,c,b],1/1)$,
$b_4=([b,a],1/1)$, $b_5=([c,b,a],1/1)$.  The first component of each ballot is a preference-ordered list of candidates, and the second is the fractional transfer value (initially set to $1/1$). The evidence consists of a \emph{header} that specifies the quota (computed according to the ACT STV instructions), the number of vacancies, and the list of competing candidates. 
The fourth line is the election result, and the remainder of the evidence are non-final states visited to compute this
outcome. Each non-final state shows uncounted ballots, tallies of candidates, piles of votes attracted, backlog of the elected, backlog of the eliminated, list of elected candidates, and list of continuing candidates.

@@

\section{The Framework Architecture}\label{sec:Arc} 
%\noindent
We progress through three macro-level phases to develop the framework. For the first phase we use HOL4 to formally specify, implement and verify the generic STV, its instantiations, verifiers, and every auxiliary assertions needed for these ends. The second step relies on the verified proof translation tool of CakeML to provably correctly translate implementations of the first phase into equivalent CakeML implementations. The last phase uses the ecosystem of CakeML and its verified compiler to generate machine executable evidence verifiers.

@@

\end{tikzpicture}
\caption{Architecture of the Framework, direction of arrow represents module dependencies}
\label{PicArc}
\end{figure}
\paragraph*{\textbf{Compilation}} 
Instantiations of CakeML's  compiler for generating machine executable verifiers happen in this module. For any deeply embedded $\mathcal{V}^{*}_{i}$, corresponding to $\hat{\mathcal{V}}_{dec}$ in \textbf{STV}$_{i}$, using proofs  established in \textbf{DeepSpec}, we instantiate the compiler to synthesise executable  verifier $\mathcal{V}_{i}$.  
\section{Generic STV Formalised}\label{sec:GenCertVer}
We proceed through three steps, namely specification, implementation and verification of implementations in HOL4 to formally verify the machine and its components.
\paragraph*{Specification}  
Recall that we require minimising the trusted computing base  for computing with generated verifiers. The specification module consists of properties which we can use to verify that our implementations of  STV algorithms in HOL4 indeed match with the description of the counting methods used.  
 In light of Definition~\ref{STVMachine} these properties   actually form the semantics of the generic STV machine.   
\paragraph*{Implementation} Our final objective for creating the
framework is to produce means for validating concrete items of evidence. To this end, we need decision procedures used for deciding whether or not a given evidence is valid. Therefore we define boolean-valued functions in the implementation sub-module to computationally realise logical declarations in the specification module. In particular, we implement decision procedures whose computational content  provably realises the specification of the machine semantics.  
\paragraph*{Verification} Here we formally prove that the implementations meet the expectations of their respective specifications. Therefore through verification we come to eliminate a trusted layer  required to lay in our framework.

@@

\begin{small}
\begin{holthmenv}
  \HOLConst{ValidInitCandList}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{distinct}\;\HOLFreeVar{l}
\end{holthmenv}
\end{small}
The predicate ValidInitCandList which realises item (6) asserts that each candidate has a pile and a tally. We also declare that elements in the first component of the tally $t$ and pile $p$ are distinct. Therefore we come to satisfy item (1).
\begin{small}
\begin{holthmenv}
  \HOLConst{Valid_PileTally}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{c}.\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{c}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{c}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})
\end{holthmenv}
\end{small}
We implement for each of the above declarations a counterpart 
%computational 
\underline{dec}ision procedure that is later translated and then extracted as part of the verifier for actual computation. To illustrate how this phase proceeds, we provide some instances. The following two functions together implement Valid\textunderscore{}PileTally.

\begin{small}
\begin{holthmenv}
  \HOLConst{Valid_PileTally_dec1}\;\HOLConst{\ensuremath{[\,]}}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{true}\\
\HOLConst{Valid_PileTally_dec1}\;(\HOLFreeVar{h}\HOLSymConst{::}\HOLFreeVar{t})\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLConst{\HOLConst{mem}}\;(\HOLConst{fst}\;\HOLFreeVar{h})\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally_dec1}\;\HOLFreeVar{t}\;\HOLFreeVar{l}
\end{holthmenv}
%\end{small}
%\vspace*{-1.6ex}
%\begin{small}
\begin{holthmenv}
  \HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{true}\\
\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;(\HOLFreeVar{l\sb{\mathrm{0}}}\HOLSymConst{::}\HOLFreeVar{ls})\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLKeyword{if}\;\HOLConst{\HOLConst{mem}}\;\HOLFreeVar{l\sb{\mathrm{0}}}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})\;\HOLKeyword{then}\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;\HOLFreeVar{ls}\;
\HOLKeyword{else}\;\HOLConst{false}
\end{holthmenv}
\end{small}
We prove that the implementations and specification match:
\begin{small}
\begin{holthmenv}
  \HOLConst{Valid_PileTally}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLConst{Valid_PileTally_dec1}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;\HOLFreeVar{l}
\end{holthmenv}
\end{small}
Conjunctions of the computational implementations  define a computational twin TransferAuxDec for the specification  of the transfer-elected semantics TransferAuxSpec.

\begin{small}
\begin{holthmenv}
  \HOLConst{TransferAuxDec}\;(\HOLFreeVar{qu}\HOLSymConst{,}\HOLFreeVar{st}\HOLSymConst{,}\HOLFreeVar{l})\;\HOLFreeVar{ba}\;\HOLFreeVar{t}\;\HOLFreeVar{t\sp{\prime}}\;\HOLFreeVar{p}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{bl}\;\HOLFreeVar{bl\sb{\mathrm{2}}}\;\HOLFreeVar{bl\sb{\mathrm{2}}\sp{\prime}}\;\HOLFreeVar{e}\;\HOLFreeVar{e\sp{\prime}}\;\HOLFreeVar{h}\;\HOLFreeVar{h\sp{\prime}}\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLConst{null}\;\HOLFreeVar{bl\sb{\mathrm{2}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{e\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{e}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{h\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{t\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{t}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{length}\;\HOLFreeVar{e}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLFreeVar{st}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{list_MEM_dec}\;(\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenDoublePlus}\;\HOLFreeVar{e})\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{list_MEM_dec}\;\HOLFreeVar{bl}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{distinct}\;(\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenDoublePlus}\;\HOLFreeVar{e})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally_dec1}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally_dec1}\;\HOLFreeVar{p}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{p}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}
\;\HOLConst{Valid_PileTally_dec1}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenNeg{}}\HOLConst{null}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{distinct}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{null}\;\HOLFreeVar{bl\sb{\mathrm{2}}\sp{\prime}}\\
\;\;\HOLConst{distinct}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{distinct}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{p})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenNeg{}}\HOLConst{null}\;\HOLFreeVar{bl}\;\HOLSymConst{\HOLTokenConj{}}
\;\HOLConst{null}\;\HOLFreeVar{ba}
\end{holthmenv}
\end{small}
